Linking hypothesis and number of response options modulate inferred scalar implicature rate

### 2018-05-10 writeup.Rmd
added abstract and conclusion and final edits'

---
title             : "Linking assumptions and number of response options modulate inferred scalar implicature rate"
shorttitle        : "Linking assumptions and implicature rate"
author:

  - name          : "Masoud Jasbi"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
  - name          : "Brandon Waldon"
    affiliation   : "1"
  - name          : "Judith Degen"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "Stanford University"
author_note: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
  Enter author note here.
abstract: |
  Enter abstract here. Each new line herein must be indented, like this line.

keywords          : "scalar implicature; methodology; linking assumption; experimental pragmatics; truth-value judgment task"
wordcount         : "X"
bibliography      : ["r-references.bib"]
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : no
class             : "man"
output            : papaja::apa6_pdf
editor_options: 
  chunk_output_type: inline
---

@@

The past 15 years have seen the rise and development of a bustling and exciting new field at the intersection of linguistics, psychology, and philosophy: *experimental pragmatics* [@Bott2004;@Breheny2006;@DegenTanenhaus2015;@Grodner2010;@huang2009;@Geurts2009;@noveck2008] **XXX ADD MORE**. Experimental pragmatics is devoted to experimentally testing theories of how language is used in context. How do listeners draw inferences about the -- often underspecified -- linguistic signal they receive from speakers? How do speakers choose between the many utterance alternatives they have at their disposal?

@@

A standard measure that has stood as a proxy for implicature rate across many studies is the proportion of 'pragmatic' judgments in truth-value judgment paradigms [@Bott2004;@Noveck2001;@Noveck2003;@Chemla2011;@Geurts2009;@DegenTanenhaus2015;@DeNeys2007;@Degen2014]. In these kinds of tasks, participants are provided a set of facts, either presented visually or via their own knowledge of the world. They are then asked to judge whether a sentence intended to describe those facts is true or false (or alternatively, whether it is right or wrong, or they are asked whether they agree or disagree with the sentence). The crucial condition for assessing implicature rates in these kinds of studies typically consists of a case where the facts are such that the stronger alternative is true and the target utterance is thus also true but underinformative. For instance, @Bott2004 asked participants to judge sentences like 'Some elephants are mammals', when world knowledge dictates that all elephants are mammals. Similarly, @DegenTanenhaus2015 asked participants to judge sentences like 'You got some of the gumballs' in situations where the visual evidence indicated that the participant received all the gumballs from a gumball machine. In these kinds of scenarios, the story goes, if a participant responds 'FALSE', that indicates that they computed a scalar implicature, eg to the effect of 'Not all elephants are mammals' or 'You didn't get all of the gumballs', which is (globally or contextually) false. If instead a participant responds 'TRUE', that is taken to indicate that they interpreted the utterance literally as `Some, and possibly all, elephants are mammals' or 'You got some, and possibly all, of the gumballs'.

Given the centrality of the theoretical notion of 'implicature rate' to much of experimental pragmatics, there is to date a surprising lack of discussion of the basic assumption that it is adequately captured by the  proportion of FALSE responses in truth-value judgment tasks (but see @BenzGotzner2014; @Geurts2009; @Degen2014; @Katsos2011). Indeed, the scalar implicature acquisition literature was shaken up when @Katsos2011 showed that simply by introducing an additional response option, children started looking much more pragmatic than had been previously observed in a binary judgment paradigm. @Katsos2011 allowed children to distribute 1, 2, or 3 strawberries to a puppet depending on 'how good the puppet said it'. The result was that children gave on average fewer strawberries to the puppet when he produced underinformative utterances compared to when he produced literally true and pragmatically felicitous utterances, suggesting that children do, in fact, display pragmatic ability even at ages when they had previously appeared not to.

But this raises an important question: in truth-value judgment tasks, how does the researcher know whether an interpretation is literal or the result of an implicature computation? The binary choice task typically used is appealing in part because it allows for a direct mapping from response options -- TRUE and FALSE -- to interpretations -- literal and pragmatic. That the seeming simplicity of this mapping is illusory becomes apparent once a third response option is introduced, as in the @Katsos2011 case. How is the researcher to interpret the intermediate option? @Katsos2011 grouped the intermediate option with the negative endpoint of the scale for the purpose of categorizing judgments as literal vs. pragmatic, i.e., they interpreted the intermediate option as pragmatic. But it seems just as plausible that they could have grouped it with the positive endpoint of the scale and taken the hard line that only truly FALSE responses constitute evidence of a full-fledged implicature. The point here is that there has been remarkably little consideration of *linking functions* between behavioral measures and theoretical constructs in experimental pragmatics, a problem in many subfields of psycholinguistics [@Tanenhaus2004]. We argue that it is time to engage more seriously with these issues.

We begin by reporting an experiment that addresses the following question: do the number of response options provided in a truth-value judgment task and the way that responses are grouped into pragmatic ('SI') and literal ('no SI') change inferences about scalar implicature rates? Note that this way of asking the question presupposes two things: first, that whatever participants are doing in a truth-value judgment task, the behavioral measure can be interpreted as providing a measure of interpretation; and second,  that listeners either do or do not compute an implicature on any given occasion. In the General Discussion we will discuss both of these issues. First, following @Degen2014, we will offer some remarks on why truth-value judgment tasks are better thought of as measuring participants' estimates of speakers' *production* probabilities. This will suggest a completely different class of linking functions. Next, we discuss an alternative conception of scalar implicature as a probabilistic phenomeonen, a view that has recently rose to prominence in the subfield of probabilistic pragmatics [@Franke2016;@Goodman2016]. This alternative conception of scalar implicature, we argue, affords developing and testing quantitative linking functions in a rigorous and motivated way.

Consider a setup in which a listener is presented a card with a depiction of either one or two animals (see Figure \@ref(fig:linkvisualization) for an example). As in a standard truth-value judgment task, the listener then observes an underinformative utterance about this card (e.g., 'There is a cat or a dog on the card') and is asked to provide a judgment on a scale with 2, 3, 4, or 5 response options, with endpoints 'wrong' and 'right'.[^1] In the binary case, this reproduces the standard truth-value judgment task. Figure \@ref(fig:linkvisualization) exemplifies (some of) the researcher's options for grouping responses. Under what we will call the 'Strong link' assumption, only the negative endpoint of the scale is interpreted as evidence for a scalar implicature having been computed. Under the 'Weak link' assumption, in contrast, any response that does not correspond to the positive endpoint of the scale is interpreted as evidence for a scalar implicature having been computed. Intermediate grouping schemes are also possible, but these are the ones we will consider here. Note that for the binary case, the Weak and Strong link return the same categorization scheme, but for any number of response options greater than 2, the Weak and Strong link can in principle lead to differences in inferences about implicature rate.

@@

Let's examine an example. Assume three response options (wrong, neither, right).  Assume further that each of the three responses was selected by a third of participants, i.e., the distributions of responses is 1/3, 1/3, and 1/3. Under the Strong link, we infer that this task yielded an implicature rate of 2/3. Under the Weak link, we infer that this task yielded an implicature rate of 1/3. This is quite a drastic difference if we are, for instance, interested in whether scalar implicatures are inference defaults and we would like to interpret an implicature rate of above an arbitrary threshold (e.g., 50%) as evidence for such a claim. Under the Strong link, we would conclude that scalar implicatures are not defaults. Under the Weak link, we would conclude that they are. In the experiment reported in the following section, we presented participants with exactly this setup. We manipulated the number of response options between participants and analyzed the results under different linking assumptions.

@@

Participants played an online card game in which they were asked to judge descriptions of the contents of cards. Different groups of participants were presented with different numbers of response options. On critical trials, participants were presented with descriptions for the cards that typically result in exhaustivity implicatures ("There is a cat on the card" when there was a cat and a dog) or scalar implicatures ("There is a cat or a dog on the card" when there was a cat and a dog). We categorized their responses on such trials according to the weak and the Strong link introduced above, and tested whether the number of response options and the linking assumptions led to different conclusions about the rate of computed implicatures in the experimental task.

@@

## Results

@@

Our primary goal in this study was to test whether the estimated implicature rate in the experimental task is affected by the linking assumption and the number of response options available to participants. To this end, we only analyzed the critical trials (exhaustive and scalar). In particular, we classified each data point from critical trials as constituting an implicature (1) or not (0) under the Strong and Weak link. Figure \@ref(fig:implicatureRatePlot) shows the resulting implicature rates by condition and link.

@@

[^3: To know more about the default priors of the "brms" package you can refer to [the brms package manual](ftp://cran.r-project.org/pub/R/web/packages/brms/brms.pdf).]

@@

We asked whether the linking assumptions and the number of response options available to participants affects the inferred implicature rate in an experimental study. The results presented here suggest they do. A linking assupmtion that considered the highest point on the scale as literal and any lower point as pragmatic (Weak link) resulted in higher implicature rates in tasks with 4 or 5 response options. A linking assumption that considered the lowest point on the scale as pragmatic and any higher point as literal (Strong link) reported lower implicature rates in tasks with 4 or 5 options. The results suggest that the choice of linking assumption is a crucial analytical step that can significantly impact the conclusions drawn from truth value judgment tasks. In particular, there is danger for pragmatic ability to be both under- and overestimated.

While the binary truth value judgement task avoids the analytic decision between Strong and Weak linking assumptions, the results reported here suggest that binary tasks can also underestimate participants' pragmatic competence. In binary tasks, participants are often given the lowest and highest points on a scale ("wrong" vs. "right") and are asked to report pragmatic infelicities using the lowest point (e.g. "wrong"). The study reported here showed that on trials with true but pragmatically infelicitous descriptions, participants often avoided the lowest point on the scale if they were given more intermediate options. Even though the option "wrong" was available to participants in all tasks, participants in tasks with intermediate options chose it less often. In computing implicature rate, this pattern manifested itself as a decrease in implicature rate under the Strong link when more response options were provided, and an increase in implicature rate under the Weak link when more response options were provided. These conclusions are in line with @Katsos2011's argument that pragmatic violations are not as severe as semantic violations and participants do not penalize them as much. Providing participants with only the extreme ends of the scale (e.g. wrong/right, false/true) when pragmatic violations are considered to be of an intermediate nature risks misrepresentation of participants' pragmatic competence. It further suggests that in studies that use binary tasks to investigate response-contingent processing, proportions of "literal" responses may be a composite of both literal and pragmatic underlying interpretations that just happen to get mapped differently onto different response options by participants.

This study did not investigate the effect of option labels on the inferred implicature rate. However, the results provided suggestive evidence that some options better capture participant intuitions of pragmatic infelicities than others. Among the intermediate options, "kinda right" was chosen most often to report pragmatic infelicities. The option "neither" was rarely used in the ternary and quinary tasks (where it was used as a midpoint), suggesting that participants interpreted pragmatic infelicities as different degrees of being "right" and not "neither right nor wrong." Therefore, options that capture degrees  of being "right" like "kinda right" may prove most suitable for capturing infelicity in the long run. We leave this as a methodological issue for future research.

The study had three further design features worth investigating in future work. First, the utterances were ostensibly produced by a blindfolded character. This was an intentional decision to control for violation of ignorance expectations with disjunction. A disjunction such as "A or B" often carries an implication or expectation that the speaker is not certain which alternative actually holds. Future work should investigate how the violation of the ignorance expectation interacts with link and number of response options in inferred implicature rate. Second, in this study we considered exhaustive and scalar implicatures with *or*. If the observed effects of link and number of response options hold in general, they should be observable using other scales, e.g., on implicatures with *some*. Finally, our experiment was designed as a guessing game and the exact goal or task-relevant Question Under Discussion of the game was left implicit. Given the past literature on QUD effects on scalar implicature, we expect that different goals -- e.g., to help the character win more points vs. to help the character be more accurate -- would affect how strict or lenient participants are with their judgments and ultimately affect implicature rate in the task [@Degen2014;@Zondervan2010]. Future work should systematically vary the goal of the game and explore its effects on the inferred implicature rate.

## Revisiting linking assumptions

On the traditional view of the link between implicature and behavior in sentence verification tasks, scalar implicature is conceptualized as a binary, categorical affair - that is, an implicature is either ‘calculated’ or it isn’t, and the behavioral reflexes of this categorical interpretation process should be straightforwardly observed in experimental paradigms. This assumption raises concerns for analyzing variation in behavior on a truth value judgment task; for example, why did the majority of respondents in the binary condition of our experiment answer “right” to an utterance of the underinformative "There is a cat or dog" when the card had both a cat and a dog on it? And why did a sizeable minority nonetheless choose "wrong" in this same condition?

To explain these data on the traditional view, we are forced to say that a) not all participants calculated the implicature; or that b) some participants who calculated the implicature did not choose the anticipated (i.e., "wrong") response due to some other cognitive process which overrode the 'correct' implicature behavior; or some mixture of (a) and (b). We might similarly posit that one or both of these factors underlie the variation in the ternary, quaternary, and quinary conditions. However, without an understanding of how to quantitatively specify the link between implicature calculation and its behavioral expression, the best we can hope for on this approach is an analysis which predicts general qualitative patterns in the data (e.g. a prediction of relatively more "right" responses than "wrong" responses in a given trial of our binary truth-value judgment task, or a prediction of a rise in the rate of response of "right"/"wrong" between two experimental conditions, given some contextual manipulation). However, we should stress that to the best of our knowledge, even a qualitative analysis of this kind of variation in behavior on sentence verification tasks - much less the effect of the number of response choices on that behavior - is largely underdeveloped scalar implicature literature.

@@

We take inspiration from the RSA approach and treat participants' behavior in our experimental tasks as the result of a soft-optimal pragmatic speaker in the RSA framework. That is, following @Degen2014, we proceed on the assumption that behavior on sentence verification tasks, such as truth value judgment tasks, is best modeled as a function of an individual’s mental representation of a cooperative interlocutor ($S_1$ in the language of RSA) rather than of a pragmatic listener who interprets utterances ($P_{L_1}$). In their paper, Degen & Goodman argue that sentence verification tasks are relatively more sensitive to contextual manipulations (such as manipulation of the Question Under Discussion) than are sentence interpretation tasks, and that this follows if sentence interpretation tasks - but not sentence verification tasks - require an additional layer of counterfactual reasoning about the intentions of a cooperative speaker.

A main desideratum of a behavioral linking hypothesis given the RSA view of pragmatic competence is to transform continuous probability distributions into categorical outputs (e.g. responses of “right”/”wrong” in the case of the binary condition of our experiment). For a given utterance $u$ and an intended communicated meaning $s$, $S_1$(u | s) outputs a conditional probability of $u$ given $s$. For example, in the binary condition of our experiment where a participant evaluated "There is a cat or a dog" when there were both animals on the card, the participant has access to the mental representation of $S_1$ and hence to the $S_1$ conditional probability of producing the utterance "cat or dog"" given a dog and cat card: $S_1$("cat or dog" | cat and dog). According to the linking hypothesis advanced here, the participant provides a particular response to $u$ if the RSA speaker probability of $u$ lies within a particular probability interval, given an observed state of the world (i.e. the configuration of animals on the card in our experiment). We model a responder, $R$, who in the binary condition responds “right” to an utterance $u$ in world $s$ just in case $S_1(u | s)$ exceeds some probability threshold $\theta$:

@@

**The following is all new** In an RSA model, $S_1$(u | s) will be defined for any possible combination of possible utterance and possible world state. One upshot of this is that for the purposes of our linking hypothesis, participants are modeled as employing the same heuristic - does $S_1$(u | s) exceed the threshold? - in both 'implicature' and 'non-implicature' conditions of a truth value judgment task experiment. That is, participants never evaluate utterances directly on the basis of logical truth or falsity: for example, our blindfolded character Bob's guess of "cat and dog" on a cat and dog card trial is "right" to the vast majority of participants not because the guess is logically true but because $S_1$("cat and dog" | cat and dog) is exceedingly high.

For further illustration, we use our definition of a pragmatically-competent speaker $S_1$ (as defined above) to calculate the speaker probabilities of utterance for "cat", "dog", "cat and dog", and "elephant" - given either a cat on the card, or both a cat and a dog on the card. In calculating these probabilities, we assume that the space of possible utterances is the set of utterances made by Bob in our experiment (i.e. any possible single, disjunctive, or conjunctive guess involving "cat", "dog", or "elephant"). For the purposes of our model, we assume a uniform cost term on all utterances. We furthermore assume that the space of possible meanings corresponds to the set of possible card configurations that a participant may have seen in our experiment, and that the prior probability distribution over these world states is uniform. Lastly, we set $\alpha$ - the speaker rationality parameter - to 1. The resulting probability values are presented in Figure 8. **not sure that this is the right figure number**

```{r message=FALSE, warning=FALSE, fig.cap = "Speaker probabilities of utterances on the exhaustive and scalar trials, as obtained using the model described in this section."}

@@

speaker_probs %>%
  ggplot(aes(x=u, y=prob)) +
  geom_bar(stat = "identity", position="dodge", width = 0.6) +
  facet_wrap(~s) +
  labs(x="", y= paste("Speaker probability ", "S(u | s)"))+
  theme_few() +
  theme(text = element_text(size=12), axis.text.x = element_text(angle = 45, hjust = 1))
  scale_fill_manual(values = c("red4", "springgreen3"), guide=FALSE)
```

>>>>

---
title             : "Linking hypothesis and number of response options modulate inferred scalar implicature rate"
shorttitle        : "Linking hypotheses and implicature rate"
author:

  - name          : "Masoud Jasbi"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "XXX Postal address"
    email         : "XXX my@email.com"
  - name          : "Brandon Waldon"
    affiliation   : "1"
  - name          : "Judith Degen"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "Stanford University"
author_note: |
  XXX Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
  Enter author note here.
abstract: |
  The past 15 years have seen increasing experimental investigations of core pragmatic questions in the ever more active and lively field of experimental pragmatics. Within experimental pragmatics, many of the core questions have relied on the operationalization of the theoretical notion of 'implicature rate'. Implicature rate based results have informed the work on acquisition, online processing, and scalar diversity, inter alia. Despite its theoretical importance. Implicature rate has  typically been quantified as the proportion of 'pragmatic' judgments in two-alternative forced choice truth value judgment tasks. Despite its theoretical importance, this linking hypothesis from implicature rate to behavioral responses has never been extensively tested. Here we show that two factors dramatically affect the 'implicature rate' inferred from truth value judgment tasks: a) the number of responses provided to participants; and b) the linking hypothesis about what constitutes a 'pragmatic' judgment. We argue that it is time for the field of experimental pragmatics to engage more seriously with its foundational assumptions about how theoretical notions map onto behaviorally measurable quantities, and present a sketch of an alternative linking hypothesis that derives behavior in truth value judgment tasks from probabilistic utterance expectations.
keywords          : "scalar implicature; methodology; linking hypothesis; experimental pragmatics; truth value judgment task"
wordcount         : "X"
bibliography      : ["r-references.bib"]
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : no
class             : "man"
output            : papaja::apa6_pdf
editor_options: 
  chunk_output_type: inline
---

@@

The past 15 years have seen the rise and development of a bustling and exciting new field at the intersection of linguistics, psychology, and philosophy: *experimental pragmatics* [@chierchia2001;@Noveck2003;@Bott2004;@Breheny2006;@DegenTanenhaus2015;@DegenTanenhaus2016;@Grodner2010;@huang2009;@Geurts2009;@noveck2008;@Bott2016;@Breheny2013;@VanTiel2014;@DeNeys2007;@Katsos2011;@Papafragou2004;@Barner2011;@Bonnefon2009;@Tomlinson2013]. Experimental pragmatics is devoted to experimentally testing theories of how language is used in context. How do listeners draw inferences about the -- often underspecified -- linguistic signal they receive from speakers? How do speakers choose between the many utterance alternatives they have at their disposal?

@@

A standard measure that has stood as a proxy for implicature rate across many studies is the proportion of 'pragmatic' judgments in truth value judgment paradigms [@Bott2004;@Noveck2001;@Noveck2003;@Chemla2011;@Geurts2009;@DegenTanenhaus2015;@DeNeys2007;@Degen2014]. In these kinds of tasks, participants are provided a set of facts, either presented visually or via their own knowledge of the world. They are then asked to judge whether a sentence intended to describe those facts is true or false (or alternatively, whether it is right or wrong, or they are asked whether they agree or disagree with the sentence). The crucial condition for assessing implicature rates in these kinds of studies typically consists of a case where the facts are such that the stronger alternative is true and the target utterance is thus also true but underinformative. For instance, @Bott2004 asked participants to judge sentences like 'Some elephants are mammals', when world knowledge dictates that all elephants are mammals. Similarly, @DegenTanenhaus2015 asked participants to judge sentences like 'You got some of the gumballs' in situations where the visual evidence indicated that the participant received all the gumballs from a gumball machine. In these kinds of scenarios, the story goes, if a participant responds 'FALSE', that indicates that they computed a scalar implicature, eg to the effect of 'Not all elephants are mammals' or 'You didn't get all of the gumballs', which is (globally or contextually) false. If instead a participant responds 'TRUE', that is taken to indicate that they interpreted the utterance literally as `Some, and possibly all, elephants are mammals' or 'You got some, and possibly all, of the gumballs'.

Given the centrality of the theoretical notion of 'implicature rate' to much of experimental pragmatics, there is to date a surprising lack of discussion of the basic assumption that it is adequately captured by the  proportion of FALSE responses in truth value judgment tasks (but see @BenzGotzner2014; @Geurts2009; @Degen2014; @Katsos2011). Indeed, the scalar implicature acquisition literature was shaken up when @Katsos2011 showed that simply by introducing an additional response option, children started looking much more pragmatic than had been previously observed in a binary judgment paradigm. @Katsos2011 allowed children to distribute 1, 2, or 3 strawberries to a puppet depending on 'how good the puppet said it'. The result was that children gave on average fewer strawberries to the puppet when he produced underinformative utterances compared to when he produced literally true and pragmatically felicitous utterances, suggesting that children do, in fact, display pragmatic ability even at ages when they had previously appeared not to.

But this raises an important question: in truth value judgment tasks, how does the researcher know whether an interpretation is literal or the result of an implicature computation? The binary choice task typically used is appealing in part because it allows for a direct mapping from response options -- TRUE and FALSE -- to interpretations -- literal and pragmatic. That the seeming simplicity of this mapping is illusory becomes apparent once a third response option is introduced, as in the @Katsos2011 case. How is the researcher to interpret the intermediate option? @Katsos2011 grouped the intermediate option with the negative endpoint of the scale for the purpose of categorizing judgments as literal vs. pragmatic, i.e., they interpreted the intermediate option as pragmatic. But it seems just as plausible that they could have grouped it with the positive endpoint of the scale and taken the hard line that only truly FALSE responses constitute evidence of a full-fledged implicature. The point here is that there has been remarkably little consideration of *linking hypothesiss* between behavioral measures and theoretical constructs in experimental pragmatics, a problem in many subfields of psycholinguistics [@Tanenhaus2004]. We argue that it is time to engage more seriously with these issues.

We begin by reporting an experiment that addresses the following question: do the number of response options provided in a truth value judgment task and the way that responses are grouped into pragmatic ('SI') and literal ('no SI') change inferences about scalar implicature rates? Note that this way of asking the question presupposes two things: first, that whatever participants are doing in a truth value judgment task, the behavioral measure can be interpreted as providing a measure of interpretation; and second,  that listeners either do or do not compute an implicature on any given occasion. In the General Discussion we will discuss both of these issues. First, following @Degen2014, we will offer some remarks on why truth value judgment tasks are better thought of as measuring participants' estimates of speakers' *production* probabilities. This will suggest a completely different class of linking hypothesiss. Next, we discuss an alternative conception of scalar implicature as a probabilistic phenomeonen, a view that has recently rose to prominence in the subfield of probabilistic pragmatics [@Franke2016;@Goodman2016]. This alternative conception of scalar implicature, we argue, affords developing and testing quantitative linking hypothesiss in a rigorous and motivated way.

Consider a setup in which a listener is presented a card with a depiction of either one or two animals (see Figure \@ref(fig:linkvisualization) for an example). As in a standard truth value judgment task, the listener then observes an underinformative utterance about this card (e.g., 'There is a cat or a dog on the card') and is asked to provide a judgment on a scale with 2, 3, 4, or 5 response options, with endpoints 'wrong' and 'right'.[^1] In the binary case, this reproduces the standard truth value judgment task. Figure \@ref(fig:linkvisualization) exemplifies (some of) the researcher's options for grouping responses. Under what we will call the 'Strong link' assumption, only the negative endpoint of the scale is interpreted as evidence for a scalar implicature having been computed. Under the 'Weak link' assumption, in contrast, any response that does not correspond to the positive endpoint of the scale is interpreted as evidence for a scalar implicature having been computed. Intermediate grouping schemes are also possible, but these are the ones we will consider here. Note that for the binary case, the Weak and Strong link return the same categorization scheme, but for any number of response options greater than 2, the Weak and Strong link can in principle lead to differences in inferences about implicature rate.

@@

Let's examine an example. Assume three response options (wrong, neither, right).  Assume further that each of the three responses was selected by a third of participants, i.e., the distributions of responses is 1/3, 1/3, and 1/3. Under the Strong link, we infer that this task yielded an implicature rate of 2/3. Under the Weak link, we infer that this task yielded an implicature rate of 1/3. This is quite a drastic difference if we are, for instance, interested in whether scalar implicatures are inference defaults and we would like to interpret an implicature rate of above an arbitrary threshold (e.g., 50%) as evidence for such a claim. Under the Strong link, we would conclude that scalar implicatures are not defaults. Under the Weak link, we would conclude that they are. In the experiment reported in the following section, we presented participants with exactly this setup. We manipulated the number of response options between participants and analyzed the results under different linking hypothesiss.

@@

Participants played an online card game in which they were asked to judge descriptions of the contents of cards. Different groups of participants were presented with different numbers of response options. On critical trials, participants were presented with descriptions for the cards that typically result in exhaustivity implicatures ("There is a cat on the card" when there was a cat and a dog) or scalar implicatures ("There is a cat or a dog on the card" when there was a cat and a dog). We categorized their responses on such trials according to the weak and the Strong link introduced above, and tested whether the number of response options and the linking hypothesiss led to different conclusions about the rate of computed implicatures in the experimental task.

@@

## Results and discussion

@@

Our primary goal in this study was to test whether the estimated implicature rate in the experimental task is affected by the linking hypothesis and the number of response options available to participants. To this end, we only analyzed the critical trials (exhaustive and scalar). In particular, we classified each data point from critical trials as constituting an implicature (1) or not (0) under the Strong and Weak link. Figure \@ref(fig:implicatureRatePlot) shows the resulting implicature rates by condition and link.

@@

[^3: For more information about the default priors of the "brms" package, see [the brms package manual](ftp://cran.r-project.org/pub/R/web/packages/brms/brms.pdf).]

@@

In this paper we asked whether linking hypothesiss and number of response options available to participants in truth value judgment tasks affects inferred implicature rates. The results presented here suggest they do. A linking assupmtion that considered the highest point on the scale literal and any lower point  pragmatic (Weak link) resulted in higher implicature rates in tasks with 4 or 5 response options compared to the standard two options. A linking hypothesis that considered the lowest point on the scale pragmatic and any higher point literal (Strong link) reported lower implicature rates in tasks with 4 or 5 options compared to the standard two options. The results suggest that the choice of linking hypothesis is a crucial analytical step that can significantly impact the conclusions drawn from truth value judgment tasks. In particular, there is danger for pragmatic ability to be both under- and overestimated.

While the binary truth value judgement task avoids the analytic decision between Strong and Weak linking hypothesiss, the results reported here suggest that binary tasks can also underestimate participants' pragmatic competence. In binary tasks, participants are often given the lowest and highest points on a scale ("wrong" vs. "right") and are asked to report pragmatic infelicities using the lowest point (e.g. "wrong"). The study reported here showed that on trials with true but pragmatically infelicitous descriptions, participants often avoided the lowest point on the scale if they were given more intermediate options. Even though the option "wrong" was available to participants in all tasks, participants in tasks with intermediate options chose it less often. In computing implicature rate, this pattern manifested itself as a decrease in implicature rate under the Strong link when more response options were provided, and an increase in implicature rate under the Weak link when more response options were provided. These observations are in line with @Katsos2011's argument that pragmatic violations are not as severe as semantic violations and participants do not penalize them as much. Providing participants with only the extreme ends of the scale (e.g. wrong/right, false/true) when pragmatic violations are considered to be of an intermediate nature risks misrepresentation of participants' pragmatic competence. It further suggests that in studies that use binary tasks to investigate response-contingent processing, proportions of "literal" responses may be a composite of both literal and pragmatic underlying interpretations that just happen to get mapped differently onto different response options by participants.

This study did not investigate the effect of response labels on the inferred implicature rate. However, the results provided suggestive evidence that some options better capture participant intuitions of pragmatic infelicities than others. Among the intermediate options, "kinda right" was chosen most often to report pragmatic infelicities. The option "neither" was rarely used in the ternary and quinary tasks (where it was used as a midpoint), suggesting that participants interpreted pragmatic infelicities as different degrees of being "right" and not "neither right nor wrong." Therefore, options that capture degrees  of being "right" like "kinda right" may prove most suitable for capturing infelicity in the long run. We leave this as a methodological issue for future research.

The study had three further design features worth investigating in future work. First, the utterances were ostensibly produced by a blindfolded character. This was an intentional decision to control for violation of ignorance expectations with disjunction. A disjunction such as "A or B" often carries an implication or expectation that the speaker is not certain which alternative actually holds. Future work should investigate how the violation of the ignorance expectation interacts with link and number of response options in inferred implicature rate. Second, in this study we considered exhaustive and scalar implicatures with *or*. If the observed effects of link and number of response options hold in general, they should be observable using other scales, e.g., on implicatures with *some*. Finally, our experiment was designed as a guessing game and the exact goal or task-relevant Question Under Discussion of the game was left implicit. Given the past literature on QUD effects on scalar implicature, we expect that different goals -- e.g., to help the character win more points vs. to help the character be more accurate -- would affect how strict or lenient participants are with their judgments and ultimately affect implicature rate in the task [@Degen2014;@Zondervan2010]. Future work should systematically vary the goal of the game and explore its effects on the inferred implicature rate. But crucially, it's unlikely that the observed effects of number of response options and linking hypothesis on inferred implicature rate are dependent on any of the discussed design choices.

## Revisiting linking hypothesiss

On the traditional view of the link between implicature and behavior in sentence verification tasks, scalar implicature is conceptualized as a binary, categorical affair -- that is, an implicature is either ‘calculated’ or it isn’t, and the behavioral reflexes of this categorical interpretation process should be straightforwardly observed in experimental paradigms. This assumption raises concerns for analyzing variation in behavior on a truth value judgment task; for example, why did the majority of respondents in the binary condition of our experiment answer “right” to an utterance of the underinformative "There is a cat or dog" when the card had both a cat and a dog on it? And why did a sizeable minority nonetheless choose "wrong" in this same condition?

To explain these data on the traditional view, we are forced to say that a) not all participants calculated the implicature; or that b) some participants who calculated the implicature did not choose the anticipated (i.e., "wrong") response due to some other cognitive process which overrode the 'correct' implicature behavior; or some mixture of (a) and (b). We might similarly posit that one or both of these factors underlie the variation in the ternary, quaternary, and quinary conditions. However, without an understanding of how to quantitatively specify the link between implicature calculation and its behavioral expression, the best we can hope for on this approach is an analysis which predicts general qualitative patterns in the data (e.g. a prediction of relatively more "right" responses than "wrong" responses in a given trial of our binary truth value judgment task, or a prediction of a rise in the rate of response of "right"/"wrong" between two experimental conditions, given some contextual manipulation). However, we should stress that to the best of our knowledge, even a qualitative analysis of this kind of variation in behavior on sentence verification tasks -- much less the effect of the number of response choices on that behavior -- is largely underdeveloped in the scalar implicature literature.

@@

We take inspiration from the RSA approach and treat participants' behavior in our experimental tasks as the result of a soft-optimal pragmatic speaker in the RSA framework. That is, following @Degen2014, we proceed on the assumption that behavior on sentence verification tasks, such as truth value judgment tasks, is best modeled as a function of an individual’s mental representation of a cooperative interlocutor ($S_1$ in the language of RSA) rather than of a pragmatic listener who interprets utterances ($P_{L_1}$). In their paper, Degen & Goodman argue that sentence verification tasks are relatively more sensitive to contextual manipulations (such as manipulation of the Question Under Discussion) than are sentence interpretation tasks, and that this follows if sentence interpretation tasks -- but not sentence verification tasks -- require an additional layer of counterfactual reasoning about the intentions of a cooperative speaker.

A main desideratum of a behavioral linking hypothesis given the RSA view of pragmatic competence is to transform continuous probability distributions into categorical outputs (e.g. responses of “right”/”wrong” in the case of the binary condition of our experiment). For a given utterance $u$ and an intended communicated meaning $s$, $S_1$(u | s) outputs a conditional probability of $u$ given $s$. For example, in the binary condition of our experiment where a participant evaluated "There is a cat or a dog" when there were both animals on the card, the participant has access to the mental representation of $S_1$ and hence to the $S_1$ conditional probability of producing the utterance "cat or dog" given a dog and cat card: $S_1$("cat or dog" | cat and dog). According to the linking hypothesis advanced here, the participant provides a particular response to $u$ if the RSA speaker probability of $u$ lies within a particular probability interval. We model a responder, $R$, who in the binary condition responds “right” to an utterance $u$ in world $s$ just in case $S_1(u | s)$ exceeds some probability threshold $\theta$:

@@

In an RSA model, $S_1$(u | s) will be defined for any possible combination of possible utterance and possible world state. One consequence of this is that for the purposes of our linking hypothesis, participants are modeled as employing the same decision criterion -- does $S_1$(u | s) exceed the threshold? -- in both 'implicature' and 'non-implicature' conditions of a truth value judgment task experiment. That is, participants never evaluate utterances directly on the basis of logical truth or falsity: for example, our blindfolded character Bob's guess of "cat and dog" on a cat and dog card trial is "right" to the vast majority of participants not because the guess is logically true but because $S_1$("cat and dog" | cat and dog) is exceedingly high.

For further illustration, we use our definition of a pragmatically-competent speaker $S_1$ (as defined above) to calculate the speaker probabilities of utterances in states of the world corresponding to our experimental conditions (i.e., for "cat", "dog", "cat and dog", and "elephant", given either a cat on the card, or both a cat and a dog on the card). In calculating these probabilities, we assume that the space of possible utterances is the set of utterances made by Bob in our experiment (i.e. any possible single, disjunctive, or conjunctive guess involving "cat", "dog", or "elephant"). For the purposes of our model, we assume a uniform cost term on all utterances. We furthermore assume that the space of possible meanings corresponds to the set of possible card configurations that a participant may have seen in our experiment, and that the prior probability distribution over these world states is uniform. Lastly, we set $\alpha$ -- the speaker rationality parameter -- to 1. The resulting speaker probabilities are shown in Figure \@ref(fig:speakerprobs).[^4]

[^4]: Note that the probabilities in each facet don't sum to 1 because the model considers all possible disjunctive, conjunctive, and simple utterances, while we are only visualizing the ones corresponding to the experimental conditions.

```{r speakerprobs, message=FALSE, warning=FALSE,fig.cap = "Speaker probabilities of utterances on the exhaustive and scalar trials, as obtained using the model described in this section."}

@@

speaker_probs %>%
  ggplot(aes(x=u, y=prob)) +
  geom_bar(stat = "identity", position="dodge", width = 0.6) +
  facet_wrap(~s) +
  labs(x="", y= paste("Speaker probability ", "S(u | s)"))+
  # theme_few() +
  theme(text = element_text(size=12), axis.text.x = element_text(angle = 45, hjust = 1))
  scale_fill_manual(values = c("red4", "springgreen3"), guide=FALSE)
```
